---
title: "Assignment 5"
author: "Kaylar Fullington"
output: pdf_document
---

## Set the working directory to the root of your DSC 520 directory
setwd("C:/Users/runek/dsc520")

## Load the `data/r4ds/heights.csv` to
```{r Make the Heights Data Frame}
heights_df <- read.csv("data/r4ds/heights.csv")
```

install.packages("Hmisc"); install.packages("ggm"); install.packages("ggplot2"); install.packages("polycor")

library(boot); library(ggm); library(ggplot2); library(Hmisc); library(polycor)

# Using `cor()` compute correlation coefficients for
## height vs. earn
```{r Corrleation Between Height and Earnings}
cor(heights_df$height, heights_df$earn, use = "everything")
```

### age vs. earn
```{r Correlation Between Age and Earnings}
cor(heights_df$age, heights_df$earn, use = "everything")
```

### ed vs. earn
```{r Corrleation Between Education and Earnings}

cor(heights_df$ed, heights_df$earn, use = "everything")
```

# Spurious correlation
The following is data on US spending on science, space, and technology in millions of today's dollars and suicides by hanging, strangulation, and suffocation for the years 1999 to 2009

## Compute the correlation between these variables
```{r Correlation between Tech Spending and Suicides}

tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)

suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)

cor(tech_spending, suicides, use = "everything")

```