---
title: "Assignment 4"
author: "Kaylar Fullington"
date: "2022-01-23"
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: "C:/Users/runek/dsc520/assignments/assignment04/bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Markdown Basics

## Favorite Foods
1. Chocolate
2. Watermelon
3. Kale

## Images


## Add a Quote

"A room without books is like a body without a soul." - Marcus Tullius Cicero

## Add an Equation
```{r simple-math-ex}
1 + 1
```
## Add a Footnote
This is a footnote[^1] in Markdown.

[^1]: Test footnote

## Add Citations

* R for Everyone [@lander2014r]
* Discovering Statistics Using R [@field2012discovering]

# Inline Code

```{r include = FALSE}
heights_df <- read.csv("C:/Users/runek/dsc520/data/r4ds/heights.csv")
covid_df <- read.csv("C:/Users/runek/dsc520/data/nytimes/covid-19-data/us-states.csv")
covid_df$date <- as.Date(covid_df$date)
date_df <- data.frame(date = covid_df$date, year = as.numeric(format(covid_df$date, format = "%Y")), month = as.numeric(format(covid_df$date, format = "%m")),
day = as.numeric(format(covid_df$date, format = "%d")))
california_df <- covid_df[ which( covid_df$state == "California"), ]
ny_df <- covid_df[which(covid_df$state == "New York"), ]
florida_df <- covid_df[which(covid_df$state == "Florida"), ]
```
## NY Times COVID-19 Data
```{r echo = FALSE}
library(ggplot2)
library(scales)
ggplot(data = florida_df, aes(x = date, group=1, color = "darkred")) +
    geom_line(aes(y = cases, color = "Florida")) +
    geom_line(data = ny_df, aes(y = cases, color = "New York")) +
    geom_line(data = california_df, aes(y = cases, color = "California")) +
    scale_colour_manual("",
                        breaks = c("New York", "California", "Florida"),
                        values = c("New York" = "darkgreen", "California" ="steelblue", "Flordia" = "darkred")) +
    xlab(" ") + ylab("Cases") + scale_y_log10()
```

## R4DS Height vs Earnings
```{r echo = FALSE}
ggplot(heights_df, aes(x = height, y = earn)) + geom_point()
```

# Tables

## Knitr Table with Kable
```{r echo = FALSE}
name <- c("Aragon", "Bilbo", "Frodo", "Galadriel", "Sam", "Gandalf", "Legolas", "Sauron", "Gollum")
race <- c("Men", "Hobbit", "Hobbit", "Elf", "Hobbit", "Maia", "Elf", "Maia", "Hobbit")
in_fellowship <- c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE)
ring_bearer <- c(FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, TRUE)
age <- c(88, 129, 51, 7000, 36, 2019, 2931, 7052, 589)

characters_df <- data.frame(name, race, in_fellowship, ring_bearer, age)
knitr::kable(head(characters_df), caption = 'One Ring to Rule Them All')
```

## Pandoc Table
```{r echo = FALSE}
name <- c("Aragon", "Bilbo", "Frodo", "Galadriel", "Sam", "Gandalf", "Legolas", "Sauron", "Gollum")
race <- c("Men", "Hobbit", "Hobbit", "Elf", "Hobbit", "Maia", "Elf", "Maia", "Hobbit")
in_fellowship <- c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE)
ring_bearer <- c(FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, TRUE)
age <- c(88, 129, 51, 7000, 36, 2019, 2931, 7052, 589)

characters_df <- data.frame(name, race, in_fellowship, ring_bearer, age)
library(pander)
pandoc.table(characters_df)
```
# References

